Election 2024

### 2024-11-03 paper/paper.qmd
reformat/update paper

# Introduction
This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris, focusing on swing states that will likely decide the election result. By aggregating high-quality polls that account for recency and sample size, we create a logistic regression model that estimates the probability of a Trump victory in each state. This approach allows us to analyze voter support patterns across the states and predict election outcomes from polling data.

@@

# Data {#sec-data}
The data used in this paper was gathered from FiveThirtyEight [@FiveThirtyEight] a website that aims to use "data and evidence to advance public knowledge". The programming language for data analysis, visualization and statistical investigation was Python [@Python] along with the packages Matplot [@Matplot], Seaborn [@Seaborn], Numpy [@Numpy], Pandas [@Pandas], Jupyter [@Jupyter], Tabulate [@Tabulate], and Sklearn [@Sklearn].

@@

# Print the table using tabulate
print(tabulate(rows, headers=['State', 'Variable', 'Mean', 'Std', 'Min', 'Max', 'Count'], tablefmt='simple_grid'))
```
**Table 1: Summary statistics of key sample variabls by state. Shows high sampl size and numeric grade among all polls of swing states.**

@@

Another aspects of the polls we wish to consider is the population size of each. Larger polls result in a stronger reduction in randomness in the response. The larger the sample size the more the response reflects the true population of the state, this is known as the law of large numbers. In table 1 we can see the variety within the states polls.
For use in the model the percentages in these polls is weighted for recenecy and size of the population.

@@

```
**Table 2: Results table for the model describd above. Show support fo trump and Kamala at the stat level accordng to avalable polls. Small standard dviation show strong results, yet the rac is very close in many states. Percentages above 50% show the favoured candidate in each state.**

In Arizona and Georgia, the model predicts the strongest likelihood of a Trump victory, with a mean polling percentage showing Trump over 2 percentage points ahead in Arizona. In Georgia, Trump's mean polling percentage is 49.39% against Harris's 47.66%.The data is thus resulting in an strong 86.85% and and 71.18% likelyhood in Trump winning in Arizona and Georgia respectively.
Trump is also predicted to win in Michigan, although by smaller margins as shown in this state. As a result the model predicts trump winning with only a 53% chance, showing his odds are near to a coin flip. With Arizona, Georgia and Michigan Trump is predicted a total of 42 of the 93 swing state votes.

@@

Kamala is predicted to win the remaining states and is given fairly high chances of locking these up. despite only winning Nevada, North Carolina and Pennsilvania by under a percntage point in the model Kamala's probablity of winning those electoral votes is very high. This is due to the small standard deviations shown in table 2 making the predicted outcome of a victory more secure.
Wisconson is shown to be her srongest and most likley victory of the swing states, with Trump only having a 15% chance here. 
Overall this givs Harris 51 of the 93 swing state votes, our model expecting a win for the Democrats and the first female US president!

## Trump overall probability
Taking The probabilities from the model above we can look at the outcomes of the 207 possible combinations of wins and loses for Trump in these swing states and deduce his overall probability of reaching 270 electoral votes. Based on the modeled probabilities his current expected votes is 41.05, below the threshold of 51 he needs based on our assumptions.

Seperating the outcomes into those that give trump the required electoral college votes, we can obtain probability that trump seizes a second term. Using probability theory, and assuming independence among the states outomes, in our approach. When accounting for all possible state outcome combinations, the model calculates that Trump has approximately a 22.5% chance of securing at least 51 electoral votes from these key swing states. This probability aggregates the likelihoods across all combinations where the sum of electoral votes meets or exceeds the 51-vote threshold. The analysis reveals that despite high individual probabilities in certain states, the overall chance of Trump amassing sufficient electoral votes from the swing states is less than one in four.

@@

The result of the model tells us that overall the polls in october are favouringh Harris to win the election. From th perpctiv of the republican party if they wish to "turn the tide" this model highlights which of the swing states are more possible to flip than others. North Carolina is the premier state predicted to go blue with the highest liklyhood of Trump winning. It would be recommended to allocate resources here to try and win a state from Harris. As far as retaining currently favoured states, Michigan is a weak position for Trump according to the model, nearing a coin flip. As this state holds 15 lectoral college votes, 4th most of the swing states, it is a key state to win. Assuming the model's correct prediction that Michigan is currently held by Trump, expect the democrats to attempt to tip the scales in their favor here.

One of the most significant insights from this analysis is the discrepancy between polling data and the betting markets. While the majority of reputable polls suggest that Kamala Harris is favored on both a national level and in most swing states, betting markets like Polymarket imply a 67% chance of Trump winning the election as of October 23, 2024 [@polymarket]. This stark disconnect raises questions about the accuracy of traditional polling methods and whether betting markets, which are financial tools driven by market forces, may offer a more precise reflection of public sentiment.

@@

Another important consideration is the possible biases in both polling and betting markets. Poll respondents may not be a representative sample of the electorate; for example, Democrats could be more likely to respond to polls, skewing the results in favor of Harris. Meanwhile, individuals who participate in betting markets might form a subset of the population that is disproportionately supportive of Trump, which could explain why the implied odds heavily favor him. Furthermore, the rapid expansion and increased liquidity of betting markets in the last few years may have improved their efficiency, making them more reflective of real-world probabilities. However, historical data still suggest that polls have been accurate 78% of the time in predicting elections [@FiveThirtyEight], indicating that the reliability of betting markets remains questionable, especially in light of their poor performance during the 2016 election, when Trump’s odds were listed at +475 (17%) just before his victory [https://www.oddsshark.com/entertainment/us-presidential-odds-2016-futures].

@@

While our model provides a structured framework for predicting the election, it has several limitations. First, the model is based on polling data available far in advance of the election, meaning there is a high degree of uncertainty, and the model may not fully capture late shifts in public opinion or external shocks (e.g., economic downturns or scandals). As polling results are based on public opinion, large shock events are likely and expected, moreover scheduled events such as debates and rallies can be expected to have large impacts on the election that cannot be modeled here. Additionally, the use of linear modeling may oversimplify the complex dynamics of voter behavior, as elections often involve nonlinear influences that are difficult to predict.

@@

Candidate Trump’s low likelihood of winning shown by our results may under-represent his support, this claim is supported by the disparity with the betting markets. This could be a result of a common problem with polling data, response rate bias. In the previous two elections, where Donald Trump was also the republican candidate, his polling was far below the eventual results on election day. Post-mortems of both elections indicated evidence of those supporting Trump not participating in polls and thus under-presenting Republican support (@clinton2021task , @kennedy2018evaluation). Should this trend continue, and be present in the polls used in this analysis, it may be expected that Trump could surprise the model and win some of these swing states. Another reason Trump could outperform the model is another bias often seen in surveys, one of social desirability. Trump is a unique candidate as, if elected, he would become the first president to be elected with federal crime conviction. This along with other negative publicity the former host of tv’s apprentice has garnered since the start of his political campaign could bias voters from publicly or privately showing their support. Even with the quality polls used in this analysis, where steps are taken to ensure representative results (see appendix B), we could be observing a skew in the results.
Another aspect that

@@

- **Demographics**: Race, age, gender, education, and income.
- **Geography**: Emphasis on swing states with representation from urban, suburban, and rural areas.
- **Voter History**: Including respondents with varying voting patterns, such as frequent voters, occasional voters, and those who are less likely to vote.

@@

- **Online Recruitment**: Targeted ads on social media, political forums, and news websites will attract a wide audience, highlighting the unique opportunity to place a small bet on the election outcome.
- **Telephone Recruitment**: To capture older demographics and those less reachable online, we will conduct phone outreach using commercial databases and voter registration information.

### Incentives for Engagement
A betting pool is a central part of this methodology, with $50,000 allocated as rewards for respondents who accurately predict the election outcome. Each participant will receive a $10 allowance to bet on their predicted winner (Trump or Harris), with payouts based on live odds, creating a direct incentive for respondents to make thoughtful, informed predictions. The purpose of this is twofold, it aims to minmize the attrition of the respondants, so that more people complete the survey, and it also will provide us insight into who people think will win rather than just who people want to win. Ideally, the purpose of this is so that people take into consideration who their friends and family, as well as their community are expecting to win.

@@

- **Cross-Referencing Survey Responses**: Survey responses will be cross-referenced with voter registration data to validate eligibility.
- **Weighting**: Data will be weighted to reflect broader population demographics, adjusting for any imbalances in representation across age, race, gender, and geographic factors. This ensures that the results better mirror the entire electorate.

@@

- **Sample Size and Recency**: More recent polls and those with larger sample sizes will be weighted more heavily.
- **Pollster Reliability**: Historical poll accuracy and transparency scores will further inform the weight of each poll.

@@

- **Recruitment**: $10,000 for targeting and engaging respondents across multiple platforms.
- **Survey Administration (Online and Phone)**: $20,000 for both online and phone-based survey collection.
- **Betting Pool**: $50,000 allocated to reward accurate predictions and enhance response quality.
- **Data Validation and Analysis**: $10,000 for verification and weighting.
- **Modeling**: $10,000 for analyzing and forecasting based on aggregated data.

## 6. Question Design and Goal Allignment
The goal of our survey is to predict who will win the election, this is the research question that guides our survey. In order to achieve this, our questions will be isolate the relevant variables by asking only one thing at a time (ceteris paribus). We will also use item specific scales over agree-disagree formats to avoid acquiescence bias. Furthermore, we will randomize response options in order to mitigate response order bias, and emphasize the surveys anonymity to minimize social desireablity bias.

@@

# Appendix B: Siena / New York Tims Pennsilvania poll

This section will discuss a particular poll from the database of polls used in the above analysis, one made by Siena College and the new york times. The New York Times partners with the Siena college in order to carry out political polls throughout the election period, these are often high quality polls according to their rating, likely because the nyt has been doing these polls for a long time. The NYT polls often emphasise a focus on the battleground states and poll then more than others making these very useful in this paper’s analysis. This action will focus on one poll done in Philadelphia, but those in other states follow the same methodology and patterns. This consistency is important for the NYT when they are reporting their results, as it makes comparison around the US states viable and reliable. In going over the methodology this appendix aims to understand what about this pollster makes them high quality and what might they leave wanting.

@@

The poll employs a response rate-adjusted stratified sampling method using the L2 voter file, which contains detailed demographic information on registered voters. Initially the interviewee's are selected randomly from a national list of registered voters, response are then weighted. The stratification, often referred to as weighing, is applied to the random sample collected in order to properly reflect the entire demographic. The stratification accounts for:
- Statehouse district
- Political party affiliation
- Race
- Gender
- Marital status
- Household size
- Turnout history
- Age
- Home ownership
The sampling is conducted in order to address telephone coverage around the state as well as nonresponse rates. Records are selected by state and sampling is separated for Pennsylvania as a whole and its major city Phyladelphia,. The voter file is then ratified and weights are based on historically modelled response ra and coverage.

## Collection
Times/ Siena polls are conducted by telephone, using live interviews. About 96% of interviews were contacted by cellphone, and interviews were conducted in both English and spanish. In the Philadelphia polls occurring between September 11th and 16th, 240,000 calls were made to nearly 118,000 likely voters (@nyt2024cross_tabs). This high call volume is key to maximising response from all groups. In a further effort to get a good sample additional effort is made to constant under-represented groups, for example people with lower education levels.

## Questionnaire design
The questions themselves included more than just the information on who a participant would vote on. Followup questions were given either because or regardless of what their previous answers were. One example would be follow up questions for those who answer as “don’t know/ refused” in order to understand some reasoning behind this answer. Those who did select an option also were asked questions to get a clear image of their sentiment or reasoning. The benefit of these additional questions is to bring context to the initial binary question.

## Methodology pros and cons
The Random contact method in which people are contacted does not provide random sampling, as non-response has made this method on it's own more redundant. (@bailey2023new). This is clearly a factor here as the response rate was below 2%. The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues (@Brick2013Nonresponse). The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population (@Kish1965Survey), Making the means generate more accurate.

With such low response rates the possibility for risk, in spite of the effort made by the pollster, is that non-response and choice are correlated, this is known as non-response bias. It is not uncommon to assume that non-response should occur at random, especially as this poll documents repeatedly calling each likely voter, however in the case of elections there is previous evidence of a bias. Research into the previous two elections, and the polls proceeding then showed a high probability that non-responders were more likely to vote for Donald Trump, who was the Republican candidate in both 2016 and 2020 (@clinton2021task, @kennedy2018evaluation). With Trump running again in 2024, his popularity being underestimated is a distinct possibility, as is mentioned in the Discussion section.

The benefit of how the survey was taken is the detail added from the additional questions asked, some key questions allow for further stratification based on the answers. For example, survey information is gathered on self-reported likelihood to vote, and obviously this gives respondents answers far more weight if they are. One documented phenomenon worth mentioning is social desirability bias, that people do are more inclined to answer with more socially acceptable answers. This could occur throughout the questionnaire, however it is very likely that respondents overestimate their likelihood to vote (@socialdesire). The pollster’s use of both modeled turnout probabilities based on historical data and self-reporting improves the accuracy of predicting who is likely to vote dramatically. The poll’s use of likely voters still brings some uncertainty to its results, as the results may be impacted by non-voters, however siena/NYT took a methodology that aims to negate this as much as possible. Another benefit of live interviews is that clarification of answers is possible and the additional depth of the data that this provides. Qualitative insights to voter's choices or indecision creates a more robust and significant dataset. By comparison this is far more valuable than a singular binary choice quiz you may see online.

In conclusion, while the pollster's methodology is robust and incorporates several best practices in survey research, inherent challenges such as low response rates and the possibility of nonresponse and social desirability biases must be acknowledged. This poll’s methodology makes great efforts to mitigate these biases through its weighting methods, yet these factors could affect the accuracy of the poll results. Nonetheless, The poll provides valuable insights and remains a useful resource for analysing voter preferences, especially due to its aforementioned weighting strategy. The poll's transparency in methodology makes it an even more reliable resource and enhances the credibility and comparability of the findings.

>>>>

# 1 Introduction
This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris, focusing on swing states that will likely decide the election result. By aggregating high-quality polls that account for recency and sample size, we create a logistic regression model that estimates the probability of a Trump victory in each state. This approach allows us to analyze voter support patterns across the states and predict election outcomes from polling data.

@@

# 2 Data {#sec-data}
The data used in this paper was gathered from FiveThirtyEight [@FiveThirtyEight] a website that aims to use "data and evidence to advance public knowledge". The programming language for data analysis, visualization and statistical investigation was Python [@Python] along with the packages Matplot [@Matplot], Seaborn [@Seaborn], Numpy [@Numpy], Pandas [@Pandas], Jupyter [@Jupyter], Tabulate [@Tabulate], and Sklearn [@Sklearn].

@@

# Print the table using tabulate
print(tabulate(rows, headers=['State', 'Variable', 'Mean', 'Std', 'Min', 'Max', 'Count'], tablefmt='simple_grid'))
```
**Figure 2: Summary statistics of key sample variables by state. Shows high sample size and numeric grade among all polls of swing states.**

@@

Another aspect of the polls we wish to consider is the population size of each. Larger polls result in a stronger reduction in randomness in the response. The larger the sample size the more the response reflects the true population of the state, this is known as the law of large numbers. In @Fig-3 we can see the variety within the state polls.
For use in the model, the percentages in these polls are weighted for recency and size of the sampled population.

@@

```
**Figure 4: Is a results table for the model described above.It shows support for Trump and Harris at the state level according to the available polls. Small standard deviations show consistent results, yet the race is very close in many states. Percentages above 50% show the favoured candidate in each state.**

In Arizona and Georgia, the model predicts the strongest likelihood of a Trump victory, with a mean polling percentage showing Trump over 2 percentage points ahead in Arizona. In Georgia, Trump's mean polling percentage is 49.39% against Harris's 47.66%. The data thus resulted in a strong 86.85% and 71.18% likelihood of Trump winning in Arizona and Georgia respectively.

Trump is also predicted to win in Michigan, although by smaller margins as shown in this state. As a result, the model predicts Trump winning with only a 53% chance, showing his odds are near to a coin flip. With Arizona, Georgia and Michigan Trump is predicted a total of 42 of the 93 swing state votes.

@@

Kamala is predicted to win the remaining states and is given fairly high chances of locking these up. despite only winning Nevada, North Carolina and Pennsylvania by under a percentage point in the model Kamala's probability of winning those electoral votes is very high. This is due to the small standard deviations shown in @Tab-2 making the predicted outcome of a victory more secure.

Wisconsin is shown to be her strongest and most likely victory of the swing states, with Trump only having a 15% chance here.

Overall this gives Harris 51 of the 93 swing state votes, our model expecting a win for the Democrats and the first female US president!

## Trump overall probability

Taking The probabilities from the model above we can look at the outcomes of the 207 possible combinations of wins and losses for Trump in these swing states and deduce his overall probability of reaching 270 electoral votes. Based on the modelled probabilities his current expected vote is 41.05, below the threshold of 51 he needs based on our assumptions.

By separating the outcomes into those that give Trump the required electoral college votes, we can obtain the probability that Trump seizes a second term, using probability theory, and assuming independence among the state outcomes in our approach. When accounting for all possible state outcome combinations, the model calculates that Trump has approximately a 22.5% chance of securing at least 51 electoral votes from these key swing states. This probability aggregates the likelihoods across all combinations where the sum of electoral votes meets or exceeds the 51-vote threshold. The analysis reveals that despite high individual probabilities in certain states, the overall chance of Trump amassing sufficient electoral votes from the swing states is less than one in four.

@@

The result of the model tells us that overall the polls in October are favouring Harris to win the election. From the perspective of the republican party if they wish to "turn the tide" this model highlights which of the swing states are more likely to flip than others. North Carolina is the premier state predicted to go blue with the highest likelihood of Trump winning. It would be recommended to allocate resources here to try and win a state from Harris. As far as retaining currently favoured states, Michigan is in a weak position for Trump according to the model, nearing a coin flip. As this state holds 15 electoral College votes, the 4th most of the swing states, it is a key state to win. Assuming the model's correct prediction that Michigan is currently held by Trump, expect the Democrats to attempt to tip the scales in their favour there.

One of the most significant insights from this analysis is the discrepancy between polling data and the betting markets. While the majority of reputable polls suggest that Kamala Harris is favoured on both a national level and in most swing states, betting markets like Polymarket imply a 67% chance of Trump winning the election as of October 23, 2024 [@polymarket]. This stark disconnect raises questions about the accuracy of traditional polling methods and whether betting markets, which are financial tools driven by market forces, may offer a more precise reflection of public sentiment.

@@

Another important consideration is the possible biases in both polling and betting markets. Poll respondents may not be a representative sample of the electorate; for example, Democrats could be more likely to respond to polls, skewing the results in favour of Harris. Meanwhile, individuals who participate in betting markets might form a subset of the population that is disproportionately supportive of Trump, which could explain why the implied odds heavily favour him. Furthermore, the rapid expansion and increased liquidity of betting markets in the last few years may have improved their efficiency, making them more reflective of real-world probabilities. However, historical data still suggest that polls have been accurate 78% of the time in predicting elections [@FiveThirtyEight], indicating that the reliability of betting markets remains questionable, especially in light of their poor performance during the 2016 election, when Trump’s odds were listed at +475 (17%) just before his victory [@oddsshark2016].

@@

While our model provides a structured framework for predicting the election, it has several limitations. First, the model is based on polling data available far in advance of the election, meaning there is a high degree of uncertainty, and the model may not fully capture late shifts in public opinion or external shocks (e.g., economic downturns or scandals). As polling results are based on public opinion, large shock events are expected, furthermore, scheduled events such as debates and rallies can be expected to have large impacts on the election that cannot be modelled here. Additionally, the use of linear modelling may oversimplify the complex dynamics of voter behaviour, as elections often involve nonlinear influences that are difficult to predict.

@@

Trump’s low likelihood of winning shown by our results may under-represent his support, this claim is supported by the disparity with the betting markets. This could be a result of a common problem with polling data, response rate bias. In the previous two elections, where Donald Trump was also the republican candidate, his polling was far below the eventual results on election day. Post-mortems of both elections indicated evidence that those supporting Trump were not participating in polls and thus under-presenting Republican support (@clinton2021task , @kennedy2018evaluation). Should this trend continue Trump could surprise the model and win some of these swing states. Another reason Trump could outperform the model is another bias often seen in surveys, one of social desirability. Trump is a unique candidate as, if elected, he would become the first president to be elected with a federal crime conviction. This along with other negative publicity the former host of TV’s Apprentice has garnered since the start of his political campaign could bias voters from publicly or privately showing their support. Even with the quality polls used in this analysis, where steps are taken to ensure representative results (see Appendix B), we could be observing a skew in the results.

@@

- **Demographics**: Race, age, gender, education, and income.

- **Geography**: Emphasis on swing states with representation from urban, suburban, and rural areas.

- **Voter History**: Including respondents with varying voting patterns, such as frequent voters, occasional voters, and those who are less likely to vote.

@@

- **Online Recruitment**: Targeted ads on social media, political forums, and news websites will attract a wide audience, highlighting the unique opportunity to place a small bet on the election outcome.

- **Telephone Recruitment**: To capture older demographics and those less reachable online, we will conduct phone outreach using commercial databases and voter registration information.

### Incentives for Engagement

A betting pool is a central part of this methodology, with $50,000 allocated as rewards for respondents who accurately predict the election outcome. Each participant will receive a $10 allowance to bet on their predicted winner (Trump or Harris), with payouts based on live odds, creating a direct incentive for respondents to make thoughtful, informed predictions. The purpose of this is twofold, it aims to minimize the attrition of the respondents, so that more people complete the survey, and it also will provide us insight into who people think will win rather than just who people want to win. Ideally, the purpose of this is so that people take into consideration who their friends, family, and their community, are expecting to win.

@@

- **Cross-Referencing Survey Responses**: Survey responses will be cross-referenced with voter registration data to validate eligibility.

- **Weighting**: Data will be weighted to reflect broader population demographics, adjusting for any imbalances in representation across age, race, gender, and geographic factors. This ensures that the results better mirror the entire electorate.

@@

- **Sample Size and Recency**: More recent polls and those with larger sample sizes will be weighted more heavily.

- **Pollster Reliability**: Historical poll accuracy and transparency scores will further inform the weight of each poll.

@@

- **Recruitment**: $10,000 for targeting and engaging respondents across multiple platforms.

- **Survey Administration (Online and Phone)**: $20,000 for both online and phone-based survey collection.

- **Betting Pool**: $50,000 allocated to reward accurate predictions and enhance response quality.

- **Data Validation and Analysis**: $10,000 for verification and weighting.

- **Modeling**: $10,000 for analyzing and forecasting based on aggregated data.

## 6. Question Design and Goal Alignment
The goal of our survey is to predict who will win the election, this is the research question that guides our survey. In order to achieve this, our questions will isolate the relevant variables by asking only one thing at a time (ceteris paribus). We will also use item-specific scales over agree-disagree formats to avoid acquiescence bias. Furthermore, we will randomize response options in order to mitigate response order bias and emphasize the survey's anonymity to minimize social desirability bias.

@@

# Appendix B: Siena / New York Tims Pennsylvania poll

This section discusses the Siena College and the New York Times poll from the database of polls used in the above analysis. The New York Times (NYT) polls often emphasize the swing states making these very useful in this paper’s analysis. This section will focus specifically on one poll done in Philadelphia, but those in other states follow the same methodology and patterns. This consistency is important for the New York Times when they are reporting their results as it makes comparisons of states reliable. In going over the methodology this appendix aims to understand what about this pollster makes them high quality and where they might improve.

@@

The poll employs a response rate-adjusted stratified sampling method using the L2 voter file, which contains detailed demographic information on registered voters. Initially, the interviewees are selected randomly from a national list of registered voters, responses are then weighed. The stratification, often referred to as weighing, is applied to the random sample collected in order to properly reflect the entire demographic. The stratification accounts for:

- Statehouse district
- Political party affiliation
- Race
- Gender
- Marital status
- Household size
- Turnout history
- Age
- Home ownership
The sampling is conducted in order to address telephone coverage around the state as well as nonresponse rates. Records are selected by state and sampling is separated for Pennsylvania as a whole and its major city Philadelphia. The voter file is then ratified and weights are based on historically modelled response rate and coverage.

## Collection
Times/Siena polls are conducted by telephone, using live interviews. About 96% of interviewees were contacted by cellphone, and interviews were conducted in both English and Spanish. In the Philadelphia polls occurring between September 11th and 16th, 240,000 calls were made to nearly 118,000 likely voters (@nyt2024cross_tabs). This high call volume is key to maximizing response from all groups. In a further effort to get a good sample additional effort is made to contact under-represented groups.

## Questionnaire design

The questions themselves included more than just the information on who a participant would vote on. Followup questions were given either because or regardless of what their previous answers were. One example would be follow-up questions for those who answer “don’t know/ refused” in order to understand some reasoning behind this answer. Those who did select an option also were asked questions to get a clear image of their sentiment or reasoning. The benefit of these additional questions is to bring context to the initial binary question.

## Methodology pros and cons

The Random contact method in which people are contacted does not provide random sampling, as non-response has made this method on its own more redundant. (@bailey2023new). This is clearly a factor here as the response rate was below 2%. The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues (@Brick2013Nonresponse). The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population (@Kish1965Survey), Making the means generated more accurate.

With such low response rates the possibility for risk, in spite of the effort made by the pollster, is that non-response and choice are correlated, this is known as non-response bias. It is not uncommon to assume that non-response should occur at random, especially as this poll documents repeatedly calling each likely voter, however in the case of elections there is previous evidence of a bias. Research into the previous two elections and the polls proceeding then showed a high probability that non-responders were more likely to vote for Donald Trump, who was the Republican candidate in both 2016 and 2020 (@clinton2021task, @kennedy2018evaluation). With Trump running again in 2024, his popularity being underestimated is a distinct possibility, as is mentioned in the discussion section.

The benefit of how the survey was taken is the detail added from the additional questions asked, some key questions allow for further stratification based on the answers. For example, survey information is gathered on self-reported likelihood to vote, and obviously this gives respondents answers far more weight if they are. One documented phenomenon worth mentioning is social desirability bias, that people are more inclined to answer with more socially acceptable answers. This could occur throughout the questionnaire, however it is very likely that respondents overestimate their likelihood to vote (@socialdesire). The pollster’s use of both modelled turnout probabilities based on historical data and self-reporting improves the accuracy of predicting who is likely to vote dramatically. The poll’s use of likely voters still brings some uncertainty to its results, as the results may be impacted by non-voters, however, Siena/NYT took a methodology that aims to negate this as much as possible. Another benefit of live interviews is that clarification of answers is possible and the additional depth of the data that this provides. Qualitative insights into voter's choices or indecision create a more robust and significant dataset. By comparison, this is far more valuable than a singular binary choice quiz you may see online.

In conclusion, while the pollster's methodology is robust and incorporates several best practices in survey research, inherent challenges such as low response rates and the possibility of nonresponse and social desirability biases must be acknowledged. This poll’s methodology makes great efforts to mitigate these biases through its weighting methods, yet these factors could affect the accuracy of the poll results. Nonetheless, The poll provides valuable insights and remains a useful resource for analyzing voter preferences, especially due to its aforementioned weighting strategy. The poll's transparency in methodology makes it an even more reliable resource and enhances the credibility and comparability of the findings.

## what is included in the poll?

## Methodology and why?
